---
title: "descriptive_stats"
author: "Rachel Kaufman"
date: "2022-12-10"
output: html_document
---
Loading in my packages...
```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(gssr)
library(dplyr)
```
Goal for this markdown: Being able to produce summary statistics in R. Visual representations, output, etc. 

Cleaning 
```{r}
data("gss_all")

gss_2018 <- gss_get_yr(2018)

gss_2018 <- gss_2018 %>% 
  select(age, occ10, marital, hapmar, wrkstat,
         childs, race, sex, reg16, agekdbrn,
         major1, major2, partyid, rincome, height, weight) %>% 
  haven::zap_labels() #removes the lables downloaded w gss data
#major2 is graduate degree field
#agekdborn is age when r had first child

gss_2018 <- gss_2018 %>% 
  mutate(female = if_else(sex == 2L, 1L, 0L))
```

For starters, let's look at the age of respondents and get some summary statistics on age. 
```{r}
head(gss_2018)

age_sex <- gss_2018 %>% 
  select(female, age) %>%
  drop_na()

age_DS <- gss_2018 %>% 
  summarise(sd = sd(age, na.rm = TRUE),
            mean = mean(age, na.rm = TRUE),
            min = min(age, na.rm = TRUE),
            max = max(age, na.rm = TRUE),
            median = median(age, na.rm = TRUE),
            range = max - min,
            IQR = IQR(age, na.rm = TRUE)
  )
```
Make sure `drop_na()` is done at the end of any pipe. Otherwise, it will drop data u will need post selecting your variables. 

Let's visualize the age demographic! 
```{r}
ggplot(data = gss_2018, aes(x = age)) + 
 geom_density(fill = "blue")

ggplot(data = gss_2018, aes(x = age)) +
geom_histogram()

```


What if I want to `group_by()` my descriptive statistics?
Let's group by sex.
```{r}

age_DS <- gss_2018 %>% 
  group_by(female) %>% 
  summarise(sd = sd(age, na.rm = TRUE),
            mean = mean(age, na.rm = TRUE),
            min = min(age, na.rm = TRUE),
            max = max(age, na.rm = TRUE),
            median = median(age, na.rm = TRUE),
            range = max - min,
            IQR = IQR(age, na.rm = TRUE)
  )
tbl_df(age_DS)
```


A prettier table. 
```{r}
install.packages("gtsummary")
library(gtsummary)
tbl_summary(age_DS)
```

What if I want to show the differences in `age` across `sex`?
```{r}
ggplot(age_sex, aes(y = age, group = female, x = female)) +
  geom_boxplot()
```
Well, very small differences. This makes since considering the previously presented summary statistics. 

Now let's look at something with likely more variety.
Let's look at `weight` differences according to `sex` using `group_by()`. 
```{r}
weight_sex <- gss_2018 %>% 
  select(female, weight, height, sex) %>% 
  drop_na()


weight_sex_DS <- weight_sex %>% 
  group_by(female) %>% 
  summarise(sd = sd(weight),
            mean = mean(weight),
            median = median(weight),
            IQR = IQR(weight))
tibble(weight_sex_DS)
tbl_summary(weight_sex_DS) 

```

Let's look at an overlapping histogram using our `female` and `weight` variables. 
```{r}

ggplot(weight_sex, aes(x = weight, color = female)) + 
  geom_histogram(color = "darkblue", fill = "lightblue", alpha = 0.6) +
  labs(title = "Weight (LB) by Female and Male") +
  facet_wrap(~female)

#OR
weight_sex <- weight_sex %>% 
  mutate(sex_factor =  if_else(sex == 2L, "Female", "Male"))

ggplot(weight_sex, aes(x = weight, fill = sex_factor)) +
  geom_histogram(alpha = .6, position = "identity", bins = 50) +
  labs(title = "Weight (LB): Female and Male",
       subtitle = "GSS 2018") +
  theme(legend.position = "bottom")
 
```

What if I wanted a box plot?
```{r}
ggplot(weight_sex, aes(x = weight, color = sex_factor)) +
  geom_boxplot() +
  labs(title = "Box Plot of Female and Male Weight (LB)",
       subtitle = "GSS 2018")
```














